\title{Aswath Python Manual}
\author{Mark Carter}
\date{\today}
\documentclass[12pt,a4paper]{article}



\usepackage{amsmath}
\usepackage[toc]{glossaries}
\usepackage{listings}
\usepackage{makeidx}
\usepackage{float}
\restylefloat{table}


\makeindex
%\thispagestyle{myheadings}
\makeglossaries
\input{glossary.ltx}


%fancy margins
%http://tex.stackexchange.com/questions/8845/how-to-move-page-number
\usepackage{fancyhdr}
\usepackage{lipsum} % only for showing some sample text
\fancyhf{} % clear all header and footers
\renewcommand{\headrulewidth}{0pt} % remove the header rule
\renewcommand{\footrulewidth}{0.1pt}
\lfoot{aswath}
\rfoot{\thepage}
\pagestyle{fancy}
\makeatletter
\let\ps@plain\ps@fancy
\makeatother


%\usepackage{threeparttable}

\begin{document}
%\pagestyle{myheadings}
\pagestyle{fancy}
\maketitle
\pagestyle{fancy}

\begin{abstract}
This article describes the 3M model that Aswath Damodaran uses in ``The Little Book of Valuation''.
\end{abstract}

\tableofcontents{}

\section{Standard Formulae}

Presented below are a number of standard formulae. It is true that

\begin{equation}
\sum\limits_{k=0}^{n-1} a r^k = a \frac{ 1-r^n} {1-r}
\label{eq_sf1}
\end{equation}

\begin{equation}
\sum\limits_{k=0}^{n} a r^k = a r \frac{ 1-r^n} {1-r}
\label{eq_sf2}
\end{equation}


Given an interest rate, $i$, (e.g. use $i=0.05$ for 5\%), then

\begin{equation} r = 1 + i, \end{equation} 
(in our example, we would set $r=1.05$) would be the correct number to use in the forumla above.

Furthermore, the present value of a perpetuity, with each payment of $c$ being received at the end of each year is given by
\begin{equation}V =\frac{c}{ i}. \end{equation} 

If $\omega$ is the dividend cover, then $\theta$, the \index{retention ratio}\textbf{retention ratio}, is given by
\begin{equation}\theta =1 - \frac{1}{ \omega}. \end{equation} 










\section{Two-stage growth model using 3M as an example}

The ultimate aim is to compute $V_c$, the value of quity of common stock.

\subsection{$g_1$, Growth rate during growth phase}

$g_1$ can be specified explcitly, or implied by the ROCE and retention ratio/di\
vidend cover. In the latter case:
\begin{equation}
g_1 = R_1  \theta _1 \equiv R_1 \left( 1- \frac{1}{\omega_1} \right).
\end{equation}





\subsection{$h_1$, Cost of capital during growth phase}

The cost of capital for the \emph{company}, \index{H1 ($h_1$)}$h_1$, which is the hurdle rate, is given by

\begin{equation}
h_1 = \frac{M_e c_e + M_d c_d}{M_E + M_d}
\label{eq_cc1a}
\end{equation}

where $c_e$ is the cost of equity, $c_d$ is the after-tax cost of debt, $M_e$ is the market value of equity and $M_d$ is the market value of debt. So $h_1$ is the weighted average of the cost of equity and debt by market value.

Now,
\begin{equation}
c_e = r_f + \beta p,
\end{equation}

where $r_f$ is the risk-free rate, $\beta$ is the \emph{unlevered} beta of the firm, and $p$ is the equity risk premium.

So, for example, if we take $r_f = 3.72\% $, $\beta=1.29$, $p=5\% $, then
\begin{equation}
c_e = 3.72\% + 1.29 \times 4 \% = 9.16 \%
\end{equation}


Damodaran uses the 10-year treasury bond rate for the \index{risk-free rate}risk-free rate. For $\beta$, he uses \emph{unlevered} betas for the sector in which the company operates. If a company operates in several sectors, then a weighted average across sectors should be used.

The cost of debt is the risk free rate, $r_f$, plus a default spread $d$, adjusted for tax:
\begin{equation}
c_d = ( r_f + d) (1 - M_t)
\end{equation}
If available, $d$ can be obtained from bond rating agencies. More likely, a rating is not available, and must be synthasised instead. Rating agencies use a variety of financial rations, but Damodaran uses interest coverage, defined as the ratio of operating income to interest expeses, to be a simple and reliable heuristic. Using the interest coverage, it is possible to look up a default spread for a similar company.

The marginal tax rate, $M_t$, is used, as opposed to the tax rate as a whole, because debt saves the company tax at the margin. For 3M, Damodaran calculates
\begin{equation}
c_d = (3.72\% + 0.75 \% ) (1- 0.40) = 2.91\%
\end{equation}
The marginal tax rate is taken for 2008, where the federal corporation tax rate is 35\%, and local taxes add 5\% to this.

Using values of $M_e = \$57b$, $M_d=\$5.3b$, and substiuting the values for $c_d$ and $c_e$ obtain above into equation (\ref{eq_cc1a}) gives
\begin{equation}
h_1 = (9.16\% \times 57 + 2.91\% \times 5.3) / (57 + 5.3) = 8.63\%
\end{equation}


\subsection{$P_1$, Present value of company during growth phase}

Let $P_1$ be the present value of the capital in the company's growth phase. Then, from the standard formula
\begin{equation}
P_1 = A_0 (1- \theta_1) r\frac{1-r^n}{1-r}
\end{equation}
where $\theta_1$ is the re-investment rate in the growth phase. Then, from our example
\begin{equation}
A_0 = 3586
\end{equation}
\begin{equation}
\theta_1 = 30\%
\end{equation}

\index{H1 ($h_1$)}
\begin{equation}
r = \frac{1+g_1}{1+h_1} = \frac{1.075}{1.0863} = 0.9896
\end{equation}
giving
\begin{equation}
P_1 = 3586 \times 0.7 \times 0.9896 (1-0.9896^5)/(1-0.9896) = 12165
\end{equation}

\subsection{$P_{t,0}$, Present value of terminal value}

Let $P_{t,6}$ be the terminal value at the end of year 5. Then 
\begin{equation}
P_{t,6} = \frac{A_6 (1 - \theta_t)}{c_t - g_t}
\label{eq15a}
\end{equation}
where $A_6$ is the after-tax operating income in year 6, $\theta_t$ is the re-investment rate in stable growth (i.e. terminal phase), $c_t= 6.7\%$ is the cost of capital in the terminal phase, $g_t=3\%$ is the stable growth rate. Now
\begin{equation}
A_6 = A_0 (1 + g_1) ^ 5 (1+ g_t) .
\end{equation}
Set
\begin{equation}
\theta_t = \frac{g_t}{c_t}
\label{eq15b}
\end{equation}
reflecting the assumption that there will be no excess returns in stable growth. In other words, the return on capital will equal the cost of capital, $c_t$.

Substitute equation (\ref{eq15b}) into (\ref{eq15a}), giving
\begin{equation}
P_{t,6} = \frac{A_6}{c_t}.
\end{equation}
Let $P_{t,0}$ denote the present value of the terminal value at the end of year 5. Thus
\begin{equation}
P_{t, 0} = \frac{P_{t,6}}{(1+h_1)^5} = \frac{A_6}{c_t (1+h_1)^5} = A_0 r^5 \frac{(1+g_t)}{c_t}
\end{equation}

In our example,
\begin{equation}
P_{t,0} = 3586 \times 0.9896^5 \times 1.03 / 0.0676 = 51856.
\end{equation}

\subsection{$V_c$, Value of equity in common stock}

Our goal is to computer $V_c$, the value of equity in common stock. We take $V_o$, the value of the operating assets, which is the PV of the discounted cashflows, i.e.
\begin{equation}
V_o = P_1 + P_{t,0}
\end{equation}
and adjust for various factors, $V_f$, the value of the company as a \emph{firm}, is given by
\begin{equation}
V_f = V_o + a_c + a_x
\end{equation}
where $a_c$ is the value of cash and marketable securities, and $a_x$ is the value of the cross-holdings in other companies.

Subtract the debt outstanding, $a_d$, to give the value of equity:
\begin{equation}
V_e = V_f - a_d.
\end{equation}
Finally, estimate a value, $a_e$, for equity options that have been granted over time to managers that are still outstanding, to give
\begin{equation}
V_c = V_e - a_e.
\end{equation}
In expanded form, this is:
\begin{equation}
\boxed{V_c = \overbrace{P_1 + P_{t,0}}^{V_o} + a_c + a_x - a_d - a_e}.
\end{equation}

In the 3M example, $a_c = 2475$, $a_x=778$, $a_d=5297$, $a_e=1216$ and $V_o = 12165 + 51856=64021$, giving
\begin{equation}
V_c = 64021 + 2475 + 778 - 5297 - 1216 = 60761.
\end{equation}


\subsection{Reconciliation}


\begin{table}[H]
\begin{tabular}{l r l}
Line item & 2007 & Per accounts\\
Non-operating income & 981m \\
Interest expense & 210m \\
Pretax Income & 6.12B & Operating income 6193m page 4 \\
Provision for Income Taxes & 1.92B  & 1964m page 40 \\
Minority Interest & 55m & 55m page 18
\end{tabular}
Per accounts consolidated income starts on page 40
\end{table}

My estimate of operating income is $6.12B + 0.210B - 0.981B = 5.349B$. Tax on operating income is therefore $\frac{5.349}{6.12} \times 1.96 = 1.713B$. My estimate of operating income after taxes is therefore $3.636B (=5.349B - 1.713B)$, cf Aswath of 3.586B \cite[p. 42]{little}.

On page 50, 
\begin{equation}
\text{FCFF}_0 = \text{ATOI} \times \text{ Reinv rate} = 3586 \times 0.7 = 2510.2
\end{equation}
From page 54, $g_1 = 7.5\%$, the growth rate of FCFF during phase 1, and $h_1 = 8.63\%$ is the cost of capital during phase 1.



\section{Blog article of 18 October 2014}

I wrote an article \cite{blog201410} that calculated upside from estimates of DCF. a listing of the code is below. The default values used were: $c_t = h_1 = 7.91\%$, $g_t = 2.95\%$, and tax rate $\tau = 21\%$.

\lstinputlisting[language=Python]{screen20141018.py}

\section{Opscale table}

\begin{table}[H]
\caption{These default values are extracted at 02 November 2014. $\tau$ is the tax rate. It is assumed that $\beta = 1$, so that $c_t = h_1 = r_f + 1 \times p$.}
\begin{tabular}{l | r}
\hline
$r_f$ & 2.95\% \\
$p$ & 4.96\% \\
$c_t$, $h_1$  & 7.91\% \\
$g_t$, $r_f$ & 2.95\% \\
$\tau$ &  21\% \\ \hline
\end{tabular}
\end{table}

\input{opscale.ltx}

Table \ref{tbl:opscale} gives the scaling that must be made to (pre-tax) operating profits to obtain the value of the common stock prior to any adjustments. The ROCE are in columns in percentages, and the rows are dividend covers.

\textbf{Example} A company has an operating profit of \pounds443m, a dividend cover of 2.2, a ROCE of 25\%, and net cash of \pounds100m. Calculate the value of capital under standard conditions.

\textbf{Solution} Looking at table \ref{tbl:opscale}, we see that the scaling factor is 15.418. So the value of capital for the company is $443 \times 15.418 + 100 = \pounds6930m$.




%\printglossary[type=main,title={Glossary},toctitle={Glossary}]
\glsaddall
\printglossaries


\begin{thebibliography}{1}

\bibitem{blog201410}
Carter, Mark.
\emph{Stocks showing at least 50\% upside using a DCF model}
Available from: $<$http://is.gd/o4tmtr$>$. [Accessed 18 October 2014].

\bibitem{dark}
Damodaran, Aswath.
\emph{The Dark Side of Valuation. Valuing young, distressed, and comples businesses}.
Pearson Education, Inc., New Jersey,
2nd Edition,
2010.

\bibitem{little}
Damodaran, Aswath.
\emph{The Little Book of Valuation. How to Value a Company, Pick a Stock, and Profit}.
John Wiley \& Sons, Inc., New Jersey,
2011
\end{thebibliography}
\printindex
\end{document}
